Domain-specific working memory loads selectively increase negative interpertations of surprised facial expressions

### 2019-10-01 Manuscript/Manuscript.tex
working on intro

Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world. In fact, facial expressions are seen across cultures ({\textbf{???}}) and some evidence suggests expressions are innate (cite). Often, facial expressions serve as a clear social signal, but this is not always the case. While a smile from a friend likely expresses a positive affective state, other cues are not so clear. For instance, a surprised facial expression could signal either a positive (e.g., winning the lottery) or negative (e.g., a snake in the woods) affective state. In the absence of a larger context, individuals differ in their tendency to interpret surprised facial expressions as either positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to both surprised faces and ambiguous scenes (Neta, Kelley, \& Whalen, 2013). This bias towards positive or negative interpretations is known as one's valence bias.
Interpreting facial expressions requires both bottom-up (e.g., perceptual input) and top-down (e.g., emotion regulation strategies) processes. A growing body of work suggests that the initial interpretation of emotionally ambiguous stimuli is negative and driven by bottom-up processes, and that arriving at a positive interpretation requires additional, top-down regulatory processes. For example, participants reliably rate surprise as negative faster than positive (Neta \& Tong, 2016). In fact, forcing participants to slow their responding during interpretations of ambiguous images shifts individuals' biases towards positivity ({\textbf{???}}). Perceptual input also contributes to valence bias. In one recent study, Neta and colleagues ({\textbf{???}}) showed that faster intial fixation, as well as longer overall fixation, on the mouth is related to more positive interpretations of surprised faces and that forcing gaze patterns to match those of modulated interpretations of surprised expressions. In short,\\
Despite the trait-like nature of this bias (Neta, Norris, \& Whalen, 2009), valence bias may be shifted, at least temporarily, by a number of experimental manipulations. As mentioned above, simple manipulations like slowing response times will shift bias ({\textbf{???}}). Additional work has shown that increases in salivary cortisol after a stressor relate to more negative interpretations of surprised faces from baseline to post-stressor (Brown, Raio, \& Neta, 2017). Similarly, participants with positive biases at baseline will interpret surprise as more negative under threat of shock ({\textbf{???}}), suggesting that taxing cognitive resources, in this case attention, reduces the ability of individuals to interpret surprised faces as positive.

Understanding the influences on decision making under ambiguous contexts sheds light on the mechanisms responsible for these individual differences. in the present study we aim to better understand how concurrent task demands (i.e., working memory load) may influence the cognitive resources used to arrive at more top-down driven interpetations of ambiguity, specifically in an emotional context.\\
Recent work suggests that ambiguity resolution in this context requires more cognitive resources/processing compared to clearly valenced faces (Mattek, Whalen, Berkowitz, \& Freeman, 2016; Neta \& Tong, 2016).
The valence bias is trait-like (Neta et al., 2009) and generalizes to non-face stimuli (Neta et al., 2013); however, it is also malleable and may differ depenending on experimental manipulations, including stress inductions or instructions to slow responding (Brown et al., 2017; Neta \& Tong, 2016). Importantly, the valence bias relates to behavior outside of the laboratory; specifically, it is known to relate to depressive symptomology ({\textbf{???}}), at least in children. Chronic negativity biases are common in numerous psychopathologies, including depression and anxiety ({\textbf{???}}).

Distractors and task irrelevant stimuli often have detrimental effects on performance in a variety of tasks {[}; cite, cite{]}. Further, domain-specific interference may further exacerbate these effects compared to domain-general stimuli (Gruber, 2001). This effect holds up in the emotional domain; for example, the Stroop task (Stroop, 1935) has been modified by some researchers to include emotional stimuli (Whalen, Bush, Shin, \& Rauch, 2006) which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Indeed, neuroimaging work supports the idea that separate systems handle attentional biasing for domain-specific (emotional vs.~non-emotional) task relevancy (Egner, Etkin, Gale, \& Hirsch, 2008).
Given that a regulatory mechanism likely contributes to positive interpretations of surprised facial expressions, domain-specific interference may cause more negative interpretations of ambiguity compared to a more domain-general interference. Mattek and colleagues (2016) recently showed that different levels of cognitive load (i.e., holding either a single or seven digit number in working memory) does not affect subjective interpretations of surprised facial expressions, but that high cognitive loads do mitigate mouse trajectories. While the authors interpret this as a distinction between trait-like biases and dynamic cognitive-motor processes, there may be more domain-specific processes (e.g., emotional components) that span across these two measures of valence bias. Given the task irrelevance of the numeric distractors in Mattek and colleagues' (2016) work, it follows that the resources required for interpreting ambiguity as positive (Neta et al., 2009) may not have been recruited for working memory maintenance, and thus no change in subjective ratings was observed.
In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

@@

The stimuli included faces from the NimStim (Tottenham et al., 2009) and Karolinska Directed Emotional Faces (Lundqvist, Flykt, \& Öhman, 1998) stimuli sets, as in previous work ({\textbf{???}}; Brown et al., 2017). The faces consisted of 34 unique identities including 11 angry, 12 happy, and 24 surprised expressions organized pseudorandomly. The scene stimuli were selected from the International Affective Picture System (Lang, Bradley, \& Cuthbert, 2008). A total of 288 scenes (72 positive, 72 negative, and 144 neutral) were selected for the image matrices. The positive and negative images did not differ on arousal (Z = -0.23, p = 0.82). The scenes were organized into low (two images) and high (six images) cognitive load of either neutral or emotional (equal number of positive and negative) images (Figure 1).

@@

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-brown_cortisol_2017}{}%
Brown, C. C., Raio, C. M., \& Neta, M. (2017). Cortisol responses enhance negative valence perception for ambiguous facial expressions. \emph{Scientific Reports}, \emph{7}(1), 15107. doi:\href{https://doi.org/10.1038/s41598-017-14846-3}{10.1038/s41598-017-14846-3}

@@

\leavevmode\hypertarget{ref-gruber_effects_2001}{}%
Gruber, O. (2001). Effects of domain-specific interference on brain activation associated with verbal working memory task performance. \emph{Cerebral Cortex (New York, N.Y.: 1991)}, \emph{11}(11), 1047--1055. doi:\href{https://doi.org/10.1093/cercor/11.11.1047}{10.1093/cercor/11.11.1047}

@@

@@

@@

@@

@@

>>>>

\hypertarget{working-memory-and-load-theory}{%
\subsection{Working memory and load theory}\label{working-memory-and-load-theory}}

Despite extensive research on the interaction of working memory and affective processes, there is much to learn concerning how cognitive and emotional processes affect one another. Executive functions, including working memory, are related to successful self-regulation, and in turn emotion regulation (Hofmann, Schmeichel, \& Baddeley, 2012). Directly comparing working memory and self-regulation of emotional responses, Schmeichel and colleagues (2008) reported that individuals with higher levels of working memory capacity demonstrated improved self-regulation towards the emotional stimuli. This suggests a connection--perhaps through some shared resource pool--between mitigated emotional responding and larger working memory resource availablility. Other work has focused on the effects of moods or affective states on working memory performance. For instance, some reports claim that both positive and negative mood interfere with working memory (Eyesenck and Calvo, 1992); however, others suggest benefits of positive mood on working memory (Yang, Yang, \& Isen, 2013). Similarly, active working memory processes may alter concurrent affective processes. For instance, actively engaging working memory can mitigate emotional responses, particularly to negative stimuli. Recent neuroimaging work reports that negative emotional responses decrease as the cognitive demands of a working memory task increase (Van Dillen, Heslenfeld, \& Koole, 2009). Additionally, following an anger induction, those with low trait rumination show faster blood pressure recovery when provided with a distractor task (Gerin, Davidson, Christenfeld, Goyal, \& Schwartz, 2006). Together, these studies suggest a resource competition between cognitive and emotional processes; in other words, when cognitive load demands are high (e.g., during active working memory maintenance), there are fewer resources available for other (i.e., affective) processes.

While previous work suggests an inhibitory effect on some emotional responses during periods with high cognitive demands, researchers have primarily focused on emotional responses to clearly valenced emotional stimuli. For instance, Schmeichel and colleagues (2008) showed participants videos intended to elicit strong negative responses (e.g., disgust) or positive responses (e.g., humor), while others have focused on comparing responses to neutral compared to negative sitmuli (Van Dillen et al., 2009). However, many emotional appraisals in day-to-day life are more nuanced than those invoked by many of the images of negative stimuli one might encounter in the lab (e.g., snakes, mutilated bodies). For example, one may appraise the content of a billboard displaying a large order of french fries as either negative or positive depending on whether or not consuming that food is (in)congruent with one's current goals. This emotional appraisal is completed under concurrent load demands--that is, the perceiver must process both the emotional stimulus (i.e., the fries) as well as actively maintain their current goal state. Maintenance of one's goal state, such as a diet, can reduce performance on some components of executive functioning tasks (Shaw \& Tiggemann, 2004). These observations are in line with Lavie and colleague's (2004) load theory, which posits that under a large cognitive load less executive resources are available to regulate incoming stimulus information.

Recently, cognitive load theory researchers have tested, and demonstrated, the domain-specificity of load and distractor interference in visual, spatial, and phonological domains (Burnham, Sabia, \& Langan, 2014). Load interference effects may also transverse other domain componenets, such as emotional compared to neutral memory content. Emotional stimuli readily capture attention compared to neutral stimuli, and this is true even in participants with amygdala damage (Hodsoll, Viding, \& Lavie, 2011; Piech et al., 2011). Given emotional stimuli's priority position in the information processing stream, it may be that cognitive loads with emotional content, compared to neutral, differentially affect concurrent emotional appraisals. Indeed, Kensigner and colleagues (2003) showed that negative emotional content slows performance on the n-back task. Other domain-specific effects have been observed in many lines of executive functions research, including those beyond the working memory domain. For example, the Stroop task (Stroop, 1935), a common measurement tool for inhibitory control, has been modified to include both emotional and non-emotional (neutral) stimuli (Whalen, Bush, Shin, \& Rauch, 2006) which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Other neuroimaging work also supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs.~non-emotional) task relevancy (Egner, Etkin, Gale, \& Hirsch, 2008). Given this evidence for dissociations of emotional and non-emotional information domains in executive functions (e.g., working memory, inhibitory control), the present work aims to clarify the interaction of emotional and non-emotional visual working memory demands on concurrent interpretations of emotional stimuli with ambiguous valence.

\hypertarget{interpreting-ambiguity}{%
\subsection{Interpreting ambiguity}\label{interpreting-ambiguity}}

Individuals differ in their tendency to interpret ambiguously valenced stimuli, like a tempting food item or a surprised facial expression, as either positive or negative. This is attributable to these stimuli's predictive value for both positive and negative outcomes. For instance, a surprised expression could signal positive (e.g., winning the lottery) or negative (e.g., a car accident) events. This affective bias is known as one's \emph{valence bias}, and a growing body of work has used both facial expressions and scenes to quantify this individual difference (Neta, Kelley, \& Whalen, 2013; Neta, Norris, \& Whalen, 2009; Neta \& Whalen, 2010). Chronic negativity biases in memory and attention are related to psychopathology, such as depression and anxiety (Mathews \& MacLeod, 2005), suggesting the importance of understanding the factors that contribute to individuals' biases. Importantly, the valence bias is a stable measure with participants showing positively correlated scores across a one year time gap (Neta et al., 2009). Despite the relative stability of the measure, experimental manipulations are capable of shifting an individuals bias (Brown, Raio, \& Neta, 2017; Neta \& Dodd, 2018; Neta et al., 2018).

Myriad factors contribute to an individual's bias, but the initial interpretation is thought to be negative across individuals. Data supporting this initial negativity hypothesis come from many studies. For instance, reaction times are faster for negative interpretations of ambiguous stimuli (Neta \& Tong, 2016). Additionally, presentation of surprised facial expressions as low spatial frequency images, which are processed more readily than high spatial frequency images, biased interpretations towards negativity (Neta \& Whalen, 2010). Under this framework, arriving at a positive interpretation requires additional, top-down regulatory processes, and there is evidence to support this as well. For example, forcing participants to slow their responding during interpretations of ambiguous images shifts individuals' biases towards positivity (Neta et al., 2018). Evidence from the neuroimaging literature supports the initial negativity hypothesis as well; more positive individuals show higher levels of BOLD activation in brain regions recruited during emotion regulation (Petro, Tong, Henley, \& Neta, 2018). Perceptual input also contributes to valence bias. In one recent study, Neta and colleagues (2017) showed that faster intial fixation on the mouth is related to more positive interpretations of surprised faces(Neta et al., 2017). Further, forcing gaze patterns to match those of the participants with the most negative or positive bias modulated interpretations of surprised expressions (Neta \& Dodd, 2018). In all, valence bias is a useful metric for understanding both trait-like components of individuals' affective biases, as well as gauging the effects of other experimental maniuplations on affective biases.

Given the evidence that a regulatory mechanism is necessary for positive interpretations of ambiguity, a demanding cognitive load might interfere with successful regulation. Indeed, Mattek and colleagues (2016) recently showed that high levels of cognitive load (i.e., holding either a single or seven digit number in working memory) mitigates mouse trajectory deviations to the modal response. However, there was no effect on subjective valence interpretations. Just as phonological and visual cognitive loads and distractors showed domain-specificity (Burnham et al., 2014), and that domain-speicificity is observed in BOLD data (Egner et al., 2008), a domain-specific cognitive load may differentially affect interpretations of ambiguity. In other words, there is likely a relationship between the qualities of stimuli held in working memory and effects on concurrent task processing. Here, we aim to test the effects of low and high working memory loads in both emotional and neutral domains on valence bias. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials, as the high cognitive demand interferes with the regulatory mechanisms used to arrive at a positive interpretation. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

@@

The stimuli included faces from the NimStim (Tottenham et al., 2009) and Karolinska Directed Emotional Faces (Lundqvist, Flykt, \& Öhman, 1998) stimuli sets, as in previous work (Brown et al., 2017; Neta \& Whalen, 2010). The faces consisted of 34 unique identities including 11 angry, 12 happy, and 24 surprised expressions organized pseudorandomly. The scene stimuli were selected from the International Affective Picture System (Lang, Bradley, \& Cuthbert, 2008). A total of 288 scenes (72 positive, 72 negative, and 144 neutral) were selected for the image matrices. The positive and negative images did not differ on arousal (Z = -0.23, p = 0.82). The scenes were organized into low (two images) and high (six images) cognitive load of either neutral or emotional (equal number of positive and negative) images (Figure 1).

@@

Increased working memory demands (i.e., a higher cognitive load) do not always result in poorer performance on concurrent tasks. For instance Baddeley -(Baddeley, 1986) reported that increasing load by adding digits to a rehearsed number did not affect accuracy on a concurrent verbal reasoning task--instead, there was an increase in the latency of response, a potential interference effect that did not alter overall accuracy.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-baddeley_working_1986}{}%
Baddeley, A. D. (1986). Working memory. \emph{Philosophical Transactions of the Royal Society of London}, \emph{302}(110), 311--324.

\leavevmode\hypertarget{ref-brown_cortisol_2017}{}%
Brown, C. C., Raio, C. M., \& Neta, M. (2017). Cortisol responses enhance negative valence perception for ambiguous facial expressions. \emph{Scientific Reports}, \emph{7}(1), 15107. doi:\href{https://doi.org/10.1038/s41598-017-14846-3}{10.1038/s41598-017-14846-3}

\leavevmode\hypertarget{ref-burnham_components_2014}{}%
Burnham, B. R., Sabia, M., \& Langan, C. (2014). Components of working memory and visual selective attention. \emph{Journal of Experimental Psychology. Human Perception and Performance}, \emph{40}(1), 391--403. doi:\href{https://doi.org/10.1037/a0033753}{10.1037/a0033753}

@@

\leavevmode\hypertarget{ref-gerin_role_2006}{}%
Gerin, W., Davidson, K. W., Christenfeld, N. J. S., Goyal, T., \& Schwartz, J. E. (2006). The role of angry rumination and distraction in blood pressure recovery from emotional arousal. \emph{Psychosomatic Medicine}, \emph{68}(1), 64--72. doi:\href{https://doi.org/10.1097/01.psy.0000195747.12404.aa}{10.1097/01.psy.0000195747.12404.aa}

\leavevmode\hypertarget{ref-hodsoll_attentional_2011}{}%
Hodsoll, S., Viding, E., \& Lavie, N. (2011). Attentional capture by irrelevant emotional distractor faces. \emph{Emotion}, \emph{11}(2), 346--353. doi:\href{https://doi.org/10.1037/a0022771}{10.1037/a0022771}

\leavevmode\hypertarget{ref-hofmann_executive_2012}{}%
Hofmann, W., Schmeichel, B. J., \& Baddeley, A. D. (2012). Executive functions and self-regulation. \emph{Trends in Cognitive Sciences}, \emph{16}(3), 174--180. doi:\href{https://doi.org/10.1016/j.tics.2012.01.006}{10.1016/j.tics.2012.01.006}

\leavevmode\hypertarget{ref-kensinger_effect_2003}{}%
Kensinger, E. A., \& Corkin, S. (2003). Effect of negative emotional content on working memory and long-term memory. \emph{Emotion}, 378--393.

@@

\leavevmode\hypertarget{ref-mathews_cognitive_2005}{}%
Mathews, A., \& MacLeod, C. (2005). Cognitive vulnerability to emotional disorders. \emph{Annual Review of Clinical Psychology}, \emph{1}, 167--195. doi:\href{https://doi.org/10.1146/annurev.clinpsy.1.102803.143916}{10.1146/annurev.clinpsy.1.102803.143916}

@@

\leavevmode\hypertarget{ref-neta_through_2018}{}%
Neta, M., \& Dodd, M. D. (2018). Through the eyes of the beholder: Simulated eye-movement experience (``SEE'') modulates valence bias in response to emotional ambiguity. \emph{Emotion}, \emph{18}(8), 1122--1127. doi:\href{https://doi.org/10.1037/emo0000421}{10.1037/emo0000421}

@@

\leavevmode\hypertarget{ref-neta_its_2018}{}%
Neta, M., Tong, T. T., \& Henley, D. J. (2018). It's a matter of time (perspectives): Shifting valence responses to emotional ambiguity. \emph{Motivation and Emotion}, \emph{42}, 258--266. doi:\href{https://doi.org/10.1007/s11031-018-9665-7}{10.1007/s11031-018-9665-7}

\leavevmode\hypertarget{ref-neta_all_2017}{}%
Neta, M., Tong, T. T., Rosen, M. L., Enersen, A., Kim, M. J., \& Dodd, M. D. (2017). All in the first glance: First fixation predicts individual differences in valence bias. \emph{Cognition \& Emotion}, \emph{31}(4), 772--780. doi:\href{https://doi.org/10.1080/02699931.2016.1152231}{10.1080/02699931.2016.1152231}

\leavevmode\hypertarget{ref-neta_primacy_2010}{}%
Neta, M., \& Whalen, P. J. (2010). The primacy of negative interpretations when resolving the valence of ambiguous facial expressions. \emph{Psychological Science}, \emph{21}(7), 901--907. doi:\href{https://doi.org/10.1177/0956797610373934}{10.1177/0956797610373934}

\leavevmode\hypertarget{ref-petro_individual_2018}{}%
Petro, N. M., Tong, T. T., Henley, D. J., \& Neta, M. (2018). Individual differences in valence bias: fMRI evidence of the initial negativity hypothesis. \emph{Social Cognitive and Affective Neuroscience}, \emph{13}(7), 687--698. doi:\href{https://doi.org/10.1093/scan/nsy049}{10.1093/scan/nsy049}

\leavevmode\hypertarget{ref-piech_attentional_2011}{}%
Piech, R. M., McHugo, M., Smith, S. D., Dukic, M. S., Van Der Meer, J., Abou-Khalil, B., \ldots{} Zald, D. H. (2011). Attentional capture by emotional stimuli is preserved in patients with amygdala lesions. \emph{Neuropsychologia}, \emph{49}(12), 3314--3319. doi:\href{https://doi.org/10.1016/j.neuropsychologia.2011.08.004}{10.1016/j.neuropsychologia.2011.08.004}

\leavevmode\hypertarget{ref-schmeichel_working_2008}{}%
Schmeichel, B. J., Volokhov, R. N., \& Demaree, H. A. (2008). Working memory capacity and the self-regulation of emotional expression and experience. \emph{Journal of Personality and Social Psychology}, \emph{95}(6), 1526--1540. doi:\href{https://doi.org/10.1037/a0013345}{10.1037/a0013345}

\leavevmode\hypertarget{ref-shaw_dieting_2004}{}%
Shaw, J., \& Tiggemann, M. (2004). Dieting and working memory: Preoccupying cognitions and the role of the articulatory control process. \emph{British Journal of Health Psychology}, \emph{9}(Pt 2), 175--185. doi:\href{https://doi.org/10.1348/135910704773891032}{10.1348/135910704773891032}

@@

\leavevmode\hypertarget{ref-van_dillen_tuning_2009}{}%
Van Dillen, L. F., Heslenfeld, D. J., \& Koole, S. L. (2009). Tuning down the emotional brain: An fMRI study of the effects of cognitive load on the processing of affective images. \emph{NeuroImage}, \emph{45}(4), 1212--1219. doi:\href{https://doi.org/10.1016/j.neuroimage.2009.01.016}{10.1016/j.neuroimage.2009.01.016}

@@

\leavevmode\hypertarget{ref-yang_positive_2013}{}%
Yang, H., Yang, S., \& Isen, A. M. (2013). Positive affect improves working memory: Implications for controlled cognitive processing. \emph{Cognition and Emotion}, \emph{27}(3), 474--482. doi:\href{https://doi.org/10.1080/02699931.2012.713325}{10.1080/02699931.2012.713325}
